Bayesian Restricted Likelihood Methods: Conditioning on Insufficient Statistics in Bayesian Regression

### 2018-08-18 revision_1/paper.tex
edits to simulation section and corresponding code

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Simulations
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation}
Data generating model:
\begin{equation}
\label{gensim2}
\begin{split}
\theta_{i}  \sim &  N(\mu, \tau^{2}),  \ i = 1, 2, \dots, 90  \\ 
y_{ij} \sim & (1-p_{i})N(\theta_{i}, \sigma^{2}) + p_{i}N(\theta_{i}, m_{i}\sigma^{2}), j = 1, 2,..., n_{i}
\end{split}
\end{equation}
with $\mu = 0, \tau^{2} = 1, \sigma^{2} = 4$. The values of $p_{i}, m_{i}$, and $n_{i}$ depend on the group and are formed using 5 replicates of the full factorial design over factors $p_{i},m_{i},n_{i}$ with levels $p_{i} = .1, .2, .3$, $m_{i} = 9, 25$, and $n_{i} = 25, 50, 100$. The full model for fitting is a corresponding normal model without contamination:
\begin{equation}
\label{fullsim2}
\begin{split}
\mu \propto 1, \ & \tau^{2} \propto \tau^{-2}, \\
\theta_{i}\sim N(\mu, \tau^{2}), \ & \sigma^{2}_{i} \sim IG(a_{s}, b_{s}),  \ i = 1, 2, \dots, 90, \\ 
y_{ij}\sim  N(\theta_{i},\sigma^{2}_{i}), \ j = 1, 2, \dots, n_{i}
\end{split}
\end{equation}
For the restricted likelihood versions we condition on robust M-estimators of location and scale in each group: $T_{i}(\by_{i}) = (\hat\theta_{i}, \hat\sigma^{2}_{i}), i = 1, 2, ..., 90$. The two versions used are Huber's and Tukey's location estimator both coupled with Huber's scale estimator. The estimates from the Bayesian models are compared to these estimates. To complete the specification of model \eqref{fullsim2} we must specify $a_{s}$ and $b_{s}$. For this we use a variety of values representing different levels of prior knowledge. For each we set $b_{s} = 4a_{s}c_{s}$ resulting in a prior mean for each $\sigma^{2}_{i}$ of $\frac{4c_{s}a_{s}}{a_{s}-1}, \ a_{s} >1$. The larger $a_{s}$, the more informative the prior. With $c_{s} = 1$ the shrinkage (for large $a_{s}$) is to the true value of $\sigma^{2} = 4$ of the variance of the good component of data in the data generating model. We consider $a_{s} = 1.25, 2.5, 5, 10, 20 $ and $c_{s} = 1/2, 1/\sqrt{2}, 1, \sqrt{2}, 2$. Values of $c_{s} \neq 1$ will result in shrinkage towards the wrong value of $\sigma^{2}$ with strength growing with $a_{s}$.

A single (\red{maybe need to do more...each run through takes about 3 hours on my laptop and current code, maybe 10 different sets would suffice}) set of data is generated from \eqref{gensim2}. The models compared are: the full model \eqref{fullsim2}, the two restricted likelihood versions of this full model, classical robust linear models fit individually to each group.

We compare fits of each model using $MSE = \frac{1}{90}\sum_{i = 1}^{90} (\hat\theta_{i} -\theta_{i})^{2}$ with $\hat\theta_{i}$ the estimate of $\theta_{i}$ from the corresponding model.

Results for MSE

\begin{figure}[H]
\centering
\includegraphics[scale = .3]{mse_sim2_facet_scale.png}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[scale = .3]{mse_sim2_facet_as.png}
\end{figure}

Big picture - restricted methods beating corresponding robust regressions (labeled rlm). MSE robustness to prior on $\sigma_{i}^{2}$

Results for KL (plots are average of the KLs computed for each group):
\begin{figure}[H]
\centering
\includegraphics[scale = .3]{kl_sim2_facet_scale.png}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[scale = .3]{kl_sim2_facet_as.png}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale = .3]{hist_sig2_ests.png}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Simulations
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Applications}
In this section we apply the restricted likelihood to both simulated and real data.

\subsection{Simulation}
For this simulation, the true data comes from the following hierarchical data generating model:
\begin{align}
\label{gensim2}
\begin{split}
& \theta_{i}  \sim   N(\mu, \tau^{2}),  \ i = 1, 2, \dots, 90  \\ 
& y_{ij} \sim (1-p_{i})N(\theta_{i}, \sigma^{2}) + p_{i}N(\theta_{i}, m_{i}\sigma^{2}), j = 1, 2,..., n_{i}
\end{split}
\end{align}
with $\mu = 0, \tau^{2} = 1, \sigma^{2} = 4$. The values of $p_{i}, m_{i}$, and $n_{i}$ depend on the group and are formed using 5 replicates of the full factorial design over factors $p_{i},m_{i},n_{i}$ with levels $p_{i} = .1, .2, .3$, $m_{i} = 9, 25$, and $n_{i} = 25, 50, 100$. Thus, the 90 groups have a varying level of outlier contamination and sample size. We wish to build models that offer good prediction for the good portion of data within each group. The full model for fitting is a corresponding normal model without contamination:
\begin{equation}
\label{fullsim2}
\begin{split}
& \mu \propto 1, \  \tau^{2} \propto \tau^{-2}, \\
& \theta_{i}\sim N(\mu, \tau^{2}), \  \sigma^{2}_{i} \sim IG(a_{s}, b_{s}),  \ i = 1, 2, \dots, 90, \\ 
& y_{ij}\sim  N(\theta_{i},\sigma^{2}_{i}), \ j = 1, 2, \dots, n_{i}
\end{split}
\end{equation}
For the restricted likelihood versions we condition on robust M-estimators of location and scale in each group: $T_{i}(\by_{i}) = (\hat\theta_{i}, \hat\sigma^{2}_{i}), i = 1, 2, ..., 90$. The two versions use Huber's and Tukey's $\psi$ function, while both versions use Huber's $\chi$ function in $\eqref{Mest}$.

To complete the specification of model \eqref{fullsim2} we must specify $a_{s}$ and $b_{s}$. For this we use a variety of values representing different levels of prior knowledge. For each we set $b_{s} = 4a_{s}c_{s}$ resulting in a prior mean for each $\sigma^{2}_{i}$ of $\frac{4c_{s}a_{s}}{a_{s}-1}, \ a_{s} >1$. The larger $a_{s}$, the more informative the prior. With $c_{s} = 1$ the shrinkage (for large $a_{s}$) is to the true value of $\sigma^{2} = 4$ of the variance of the good component of data in the data generating model. We consider $a_{s} = 1.25,  5, 10,$ and $c_{s} = 1/2, 1, 2$. Values of $c_{s} \neq 1$ will result in shrinkage towards the wrong value of $\sigma^{2}$.

For each simulated data set, the Bayesian models are fit using MCMC. The MCMC for the restricted likelihood version requires no further computational details other than those described for the traditional Bayesian model in Section \ref{BayesLinMod}. This is because there are conditioning statistics for each group and the model's conditional independence between the groups allows the data augmentation described earlier to be performed independently within each group. That is, there is a separate Gibbs step for each group  generating group level data matching the statistics for that group.

The models are compared using Kullback-Leibler (KL) divergence from the distribution of good data to the posterior predictive distribution in order to assess the predictive capability of the models for the good part of the data. Specifically, the $i^{th}$ group of the $k^{th}$ simulated data set $\by_{k}$ compute:
\begin{equation}
\label{kl}
KL^{(M)}_{ik} = \int \log \frac{f(\tilde y | \theta_{i}, \sigma^{2})}{f_{i}(\tilde y | M, \by_{k})}  f(\tilde y | \theta_{i}, \sigma^{2}) \ dy
\end{equation}
where $M$ indexes the fitting model, $f(\tilde y | \theta_{i}, \sigma^{2})$ and is the pdf of a $N(\theta_{i}, \sigma^{2})$ at $\tilde y$ with $\sigma^{2} = 4$. For the Bayesian models ${f_{i}(\tilde y | M, \by_{k})} = \int f(\tilde y |\theta_{i}, \sigma_{i}^{2}) \pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})$ where $ f(\tilde y |\theta_{i}, \sigma_{i}^{2})$ is again a normal pdf and  $\pi(\theta_{i}, \sigma^{2}_{i} | M, \by_{k})$ is the posterior for the $i^{th}$ group model parameters under model $M$ for the $k^{th}$ data set. $M$ denotes either the full normal theory model \eqref{fullsim2} or one of the two restricted likelihood versions discussed above, where conditioning is only on the summary statistics (not the complete $\by_{k}$). For the classical robust fits, we set $f_{i}(\tilde y | M, \by_{k})$ a $N(\hat\theta_{i}, \hat\sigma^{2}_{i})$ at  $\tilde y$ as a plug in estimator for the predictive distribution. The overall mean $\overline{KL}^{(M)}_{{\cdot}{\cdot}} = \frac{1}{90K} \sum_{k = 1}^{K} \sum_{i=1}^{90} KL^{(M)}_{ik}$ is used to compare the models with $SE(\overline{KL}^{(M)}_{{\cdot} k}) = \sqrt{\frac{K}{K-1}\sum_{k = 1}^{K} (\overline{KL}^{(M)}_{{\cdot} k} - \overline{KL}^{(M)}_{{\cdot}{\cdot}})^{2}}$ used to summarize sampling variation.

Figure \ref{} displays the results for each $a_{s} = 1.25,  5, 10,$ and $c_{s} = 1/2, 1, 2$ over $K = ?? $ data sets.

\begin{figure}[H]
\centering
\includegraphics[width = 6in, height = 4in]{kl_sim2_facet_scale.png}
\end{figure}
%
%\begin{figure}[H]
%\centering
%\includegraphics[scale = .3]{kl_sim2_facet_as.png}
%\end{figure}

@@

%\begin{figure}[H]
%\centering
%\includegraphics[scale = .3]{hist_sig2_ests.png}
%\end{figure}
